### 2024-06-01 README.Rmd
More writing

In a sense, an AI is like a crutch. It can help someone to accomplish a task more easily, but it will not help if used constantly. However, what the public is being promised is not a crutch: it is an All-Terrain Vehicle that the end user cannot hope to master^[Reports of people getting their fingers crushed by Tesla's cybertruck could be a result of 1) people trusting the technology more than they should; and 2) not understanding the technology: apparently the algorithm uses increasing force to close the trunk if it fails to latch on back-to-back cycles. Hence, the trunk will not slash a banana on its first attemt to latch, but will crush a finger on the second attempt [@lambert_tesla_2024].] due to its blackbox nature (in the sense that neural networks are blackboxes, and the development of AIs itself is a blackbox where things change based on the interests of their corporate developers^[Sometimes said interest may be relatively benign, like avoiding reputational damage [by correcting a wrong](https://www.bbc.com/news/technology-65202597), but there is no guarantee that corporate interests will always align with those of the public, and in fact, they seldom will as long as there are only a handful of American companies that offer an "essential" service.], and not the interests of their relatively small customers^[And this refers not only to individuals playing with ChatGPT and MidJourney. In 2022, McMaster University operated on revenues of less than [\$843 million](https://financial-affairs.mcmaster.ca/app/uploads/2023/06/2023-24-Consolidated-Budget-FINAL.pdf) (CAD). Microsoft, in contrast, operated on revenues of approximately [\$198,000 millions](https://www.microsoft.com/investor/reports/ar22/) (USD) in 2022. To put things in perspective (and using the average exchange rate of 2022), if McMaster used all of its 2022 revenue to pay Microsoft, that would leave McMaster with zero money and still account for less than _one third of one percent point_ of Microsoft's revenue. A mid-sized Canadian university is puny compared to the economic might of Microsoft.] who fancy themselves creators but never learned (or forgot) how to code.).

@@

@@

@@

When using a Generative AI we must ask whose perspectives the tool will tend to reproduce and amplify. In this sense, numerous observers [e.g., @ahuja_we_2023; @stross_tech_2023; @gebru_tescreal_2024] have noted the ideological leanings of various actors in the AI scene, an elite world of tech investors^[A group that author Charles Stross colorfully [characterizes](https://www.antipope.org/charlie/blog-static/2024/04/the-radiant-future-of-1995.html) as "fascist-adjacent straight white males with an unadmitted political agenda". This characterization is somewhat unfair: both [Sam Altman](https://www.advocate.com/news/sam-altman-openai-exclusive) and [Peter Thiel](https://www.npr.org/2016/07/21/486966882/i-am-proud-to-be-gay-tech-investor-peter-thiel-tells-gop-convention) are openly gay.] known for fringe ideologies that combine a peculiar mix of uber-techno-optimism and existential fright (including transhumanism, extropianism, singularitarianism, and cosmism)^[A quick summary of these believes is that humans (or at lest some of them, those who can afford to do so) are destined to use technology to indefinitely extend life and conquer the stars, but not before super-intelligent AIs destroy humanity.], paired with extreme theories of justice (including rationalism, effective altruism^[A movement exposed to the harsh light of public opinion for being the philosophy espoused by convicted fraudster and money launderer Sam Bankman-Fried [@crook_effective_2023].], and longtermism^[A drastic form of utilitarianism convinced that the wellbeing of trillions of humans in the future is well worth the sacrifices of a few billions of humans today. Curiously, the contribution of the titans of technology promoting AI is not to sacrifice themselves: as evidence of the abuses inflicted in the name of innovation continues to pile up, it is clear that some of us will suffer and die--but that is a sacrifice that wealthy and well connected longtermists are willing to make.]).

@@

Search engines were already known to reinforce racism [@noble_algorithms_2018]; given the above, it really is too much to expect that AI-augmented search engines will do better^[A report by the Center for Countering Digital Hate found that Google's Bard generated wrong and/or hateful outputs in 78 out of 100 tests [@center_for_countering_digital_hate_googles_2023]. In addition to being misleading to the unadvised, AI is also a fantastic tool ripe for abuse by those who are advised _and_ have an agenda. Google's own researchers have come to the inevitable conclusion that it is easier to produce disinformation with AI, than it is to use AI to combat misinformation [@maiberg__google_2024].]. In another wretched incident, Google's AI overview stated that Barack Obama was the first Muslim president of the US, a regrettable piece of misinformation that fails to surprise given the years-long Russian-assisted right-wing propaganda campaign to otherize Obama [@tenbarge_glue_2024]. Mishaps like this, and others, are dismissed by Google as attempts to "to trip up the technology with uncommon questions" [@tenbarge_glue_2024]. This dismissal, though, only reinforces the notion that AIs are only ever useful when [regressing to the mean](#really-what-do-they-do). Responses like fiddling with the model (fine tuning it) to provide more palatable outputs (like in the case of **FIND EXAMPLE**) are not more reassuring. Who is to say what a handful of unaccountable corporations will find palatable tomorrow, let alone in the face of powerful authoritarian regimes?

@@

I suggest that universities should avoid the mistakes of the media, an industry that has innovated itself [out of the trust of the public](https://reutersinstitute.politics.ox.ac.uk/news/julia-angwin-fears-public-sphere-about-get-worse-ai-makes-it-easier-flood-zone-misinformation) (also see the "pivot-to-video" and how tech companies suckered many media companies, including ). Tech corporations (or least ways some of them), can still hope to recoup some of their investment in AI. But a university gambles its most important asset when betting on AI: its reputation as a place were knowledge is created, and as a custodian of knowledge. AIs already have a reputation for producing mis- and disinformation--from the very titans of the field^[Google could not care less about its reputation. It is [too big to care](https://pluralistic.net/2024/04/04/teach-me-how-to-shruggie/#kagi).] to the unaware^[], and beyond, to bad-faith actors. In my view, there are no efficiencies, no promises of future excellence, that can compensate the risk of ruining an institutions reputation.

>>>>

In a sense, an AI is like a crutch. It can help someone to accomplish a task more easily, but it will not help if used constantly. However, what the public is being promised is not a crutch: it is an All-Terrain Vehicle that the end user cannot hope to master^[Reports of people getting their fingers crushed by Tesla's cybertruck could be a result of 1) people trusting the technology more than they should; and 2) not understanding the technology: apparently the algorithm uses increasing force to close the trunk if it fails to latch on back-to-back cycles. Hence, the trunk will not slash a banana on its first attemt to latch, but will crush a finger on the second attempt [@lambert_tesla_2024].] due to its blackbox nature (in the sense that neural networks are blackboxes, and the development of AIs itself is a blackbox where things change based on the interests of their corporate developers^[Sometimes said interest may be relatively benign, like avoiding reputational damage [by correcting a wrong](https://www.bbc.com/news/technology-65202597), but there is no guarantee that corporate interests will always align with those of the public, and in fact, they seldom will as long as there are only a handful of American companies that offer an "essential" service.], and not the interests of their relatively small customers^[And this refers not only to individuals whose skills are yoked to ChatGPT and MidJourney. In 2022, McMaster University operated on revenues of less than [\$843 million](https://financial-affairs.mcmaster.ca/app/uploads/2023/06/2023-24-Consolidated-Budget-FINAL.pdf) (CAD). Microsoft, in contrast, operated on revenues of approximately [\$198,000 millions](https://www.microsoft.com/investor/reports/ar22/) (USD) in 2022. To put things in perspective (and using the average exchange rate of 2022), if McMaster used _all_ of its 2022 revenue to pay Microsoft, that would leave McMaster with zero money and still account for less than _one third of one percent point_ of Microsoft's revenue. A mid-sized Canadian university is like an ant compared to the economic boot of Microsoft.] who fancy themselves creators but never learned (or forgot) how to code.).

@@

Tools to posion AI models:

@salman_raising_2023

@shan_nightshade_2024

@xu_shadowcast_2024

@@

https://www.theguardian.com/commentisfree/article/2024/may/30/ugly-truth-ai-chatgpt-guzzling-resources-environment

@@

When using a Generative AI we must ask whose perspectives the tool will tend to reproduce and amplify. In this sense, numerous observers [e.g., @ahuja_we_2023; @stross_tech_2023; @gebru_tescreal_2024] have noted the ideological leanings of various actors in the AI scene, an elite world of tech investors^[A group that author Charles Stross colorfully [characterizes](https://www.antipope.org/charlie/blog-static/2024/04/the-radiant-future-of-1995.html) as "fascist-adjacent straight white males with an unadmitted political agenda". This characterization is somewhat unfair: both [Sam Altman](https://www.advocate.com/news/sam-altman-openai-exclusive) and [Peter Thiel](https://www.npr.org/2016/07/21/486966882/i-am-proud-to-be-gay-tech-investor-peter-thiel-tells-gop-convention) are openly gay. On the other hand, their political agenda hides in plain sight. @thiel_education_2009 has written about how he "no longer believe[s] that freedom and democracy are compatible". Elon Musk, well, Elon Musk once shared a meme comparing Canada's PM Justin Trudeau to Hitler, when defending the vandals who [terrorized Ottawa](https://fortune.com/2022/01/27/elon-musk-twitter-freedom-convoy-canadian-truckers-rule-covid-vaccine/) in 2022.] known for fringe ideologies that combine a peculiar mix of uber-techno-optimism and existential fright (including transhumanism, extropianism, singularitarianism, and cosmism)^[A quick summary of these beliefs is that humans (or at lest some of them, those who can afford to do so) are destined to use technology to indefinitely extend life and conquer the stars...just not before super-intelligent AIs destroy humanity.], paired with extreme theories of justice (including rationalism, effective altruism^[A movement exposed to the harsh light of public opinion for being the philosophy espoused by convicted fraudster and money launderer Sam Bankman-Fried [@crook_effective_2023].], and longtermism^[A drastic form of utilitarianism convinced that the wellbeing of trillions of humans in the future is well worth the sacrifices of a few billions of humans today. Curiously, the contribution of the titans of technology promoting AI is not to sacrifice themselves: as evidence of the abuses inflicted in the name of innovation continues to pile up, it is clear that some of us will suffer and die--but that is a sacrifice that wealthy and well connected longtermists are willing to make.]).

@@

Search engines were already known to reinforce racism [@noble_algorithms_2018]; given the above, it really is too much to expect that AI-augmented search engines will do better^[A report by the Center for Countering Digital Hate found that Google's Bard generated wrong and/or hateful outputs in 78 out of 100 tests [@center_for_countering_digital_hate_googles_2023]. In addition to being misleading to the unadvised, AI is also a fantastic tool ripe for abuse by those who are advised _and_ have an agenda. Google's own researchers have come to the inevitable conclusion that it is easier to produce disinformation with AI, than it is to use AI to combat misinformation [@maiberg__google_2024].]. In another wretched incident, Google's AI overview stated that Barack Obama was the first Muslim president of the US, a regrettable piece of misinformation that fails to surprise given the years-long Russian-assisted right-wing propaganda campaign to otherize Obama [@tenbarge_glue_2024]. Mishaps like this, and others, are dismissed by Google as attempts to "to trip up the technology with uncommon questions" [@tenbarge_glue_2024]. This dismissal, though, only reinforces the notion that AIs are only ever useful when [regressing to the mean](#really-what-do-they-do). Responses like fiddling with the model (fine tuning it) to provide more palatable outputs^[Like in the case of [Google's response](https://blog.google/products/search/ai-overviews-update-may-2024/) to the poor quality of its recently launched AI overviews.] are not more reassuring. Who is to say what a handful of unaccountable corporations will find palatable tomorrow, let alone in the face of powerful authoritarian regimes?

@@

I suggest that universities should avoid the mistakes of the media, an industry that has innovated itself [out of the trust of the public](https://reutersinstitute.politics.ox.ac.uk/news/julia-angwin-fears-public-sphere-about-get-worse-ai-makes-it-easier-flood-zone-misinformation) (also see the "pivot-to-video" and how tech companies suckered many media companies, including ). Tech corporations (or least ways some of them), can still hope to recoup some of their investment in AI. But a university gambles its most important asset when betting on AI: its reputation as a place were knowledge is created, and as a custodian of knowledge. AIs already have a reputation for producing mis- and disinformation--from the very titans of the field^[Entities like Google and Amazon could not care less about their reputation. They are [too big](https://www.businessinsider.com/ex-amazon-ghaderi-exec-suing-ai-race-copyright-allegations-2024) [to care](https://pluralistic.net/2024/04/04/teach-me-how-to-shruggie/).] to the unaware^[Like a B.C. lawyer who was [reprimanded](https://www.cbc.ca/news/canada/british-columbia/lawyer-chatgpt-fake-precedent-1.7126393) for using AI hallucinations in a legal filing. This was not even the first time that supposedly highly skilled professionals were castigated for [using](https://www.washingtonpost.com/technology/2023/11/16/chatgpt-lawyer-fired-ai/) [chatbots](https://www.reuters.com/legal/new-york-lawyers-sanctioned-using-fake-chatgpt-cases-legal-brief-2023-06-22/).], and beyond, to bad-faith actors^[Like rival geopolitical powers that are relatively impervious to propaganda themselves for reason of being [authoritarian countries](https://www.theatlantic.com/magazine/archive/2024/06/china-russia-republican-party-relations/678271/).]. In my view, there are no efficiencies, no promises of future excellence, that can compensate the risk of ruining an institutions reputation.
